\LARGE Untangling the link between traits, size and growth rate in plants

### 2016-10-18 ms/ms.tex
Incorporate suggestions from Georges Kunstler
also addresses some issues raised by Remko

% XX - The main comment I have is whether the strength of the 'correlation' between traits and growth rates can be summarized in some metric, so that it can be directly shown. To me Fig 4 is not very clear on showing how size affects the trait-growth correlations. For example, for wood density, I see a relation for all size classes. For the largest trees, it may look like a weaker correlation but isn't that because growth rates are lower? The same with Narea, to me they all look the same (except again, larger trees grow slower but it’s hard to tell there for me whether the trait-growth correlation is any different)
% XX -  The shade tolerance (WPLCP) analysis seems off-topic (it also not mentioned in the abstract). Can this be integrated more closely with the main objective, i.e. correlation between traits and growth rate and its dependence on size?

@@

Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citet{Gibert-2016}, we lacked any clear theoretical expectations on why the effect of traits may be moderated by size. Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}. A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to LMA \citep{Lambers-1992, Cornelissen-1996, Wright-2000}. An extension of the model suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}. But as noted above, the prediction for large plants is not supported by empirical results. Meanwhile, theoretical predictions on how other traits should influence growth are largely absent. Without further guidance, many researchers expected traits to linearly map onto growth rates, i.e. there is a ``fast'' and ``slow'' growth end for each trait \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.

One possible reason that theoretical predictions have not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) or mass production \citep{Wright-2000,Enquist-2007}. By contrast, the physiology of traits such as LMA and wood density suggests they have influence by altering the allocation of biomass among different tissues, rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}. A second concern for theory focussed on mass production is that measuring mass production is really only practical for small plants that can be easily harvested. For larger plants, growth is more often measured as increment in either diameter or height. For example, stem diameter growth has been measured for many decades across many thousands of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}. Theory considering only primary production therefore offers limited insight.

Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle. Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}. Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. As noted above, the paper by \citet{Gibert-2016} outlined a framework describing how traits influence growth. Here we extend the results of \citet{Gibert-2016} to provide a functioning model and explain, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size. We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment. Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}. In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}, and all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance also varies among species, correlates with traits \citep{Lusk-2008,Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Lusk-2008}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.

@@

The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig. \ref{fig:conceptual}a). We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=live (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h). The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$. A summary of all variables, units and definitions is given in Table \ref{tab:definitions}-\ref{tab:traits}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.

We assume growth is fundamentally driven by biomass production and its subsequent distribution throughout the plant. Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000}:
\begin{equation}\label{eq:dbdt}
\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
= \underbrace{\strut \alpha_{\rm bio} }_\textrm{mass per C} \,
\underbrace{\strut \alpha_{\rm y} }_\textrm{yield}
\big( \underbrace{\strut A_{\rm l} \, \bar{p}(E)}_\textrm{photosynthesis} -
\underbrace{\strut \, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}}_\textrm{respiration}\big)
- \underbrace{\strut \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}}_\textrm{turnover}.
\end{equation}
Photosynthesis is the product of the average photosynthetic rate per unit leaf area ($\bar{p}(E)$) and total leaf area ($A_{\rm l}$). We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (Fig. \ref{fig:wplcp_idea}; for details, see Supplementary Materials), and respiration and turnover rates of different tissues are constants that might differ with traits (see below). The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for growth respiration and the conversion of CO$_2$ into units of biomass, respectively. While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment, so that we can better understand the intrinsic trait- and size-related effects. Also, many vegetation models use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g. as functions of temperature. Such detail is not needed here because it will not change general model behaviour, even it would alter  he absolute values of predicted growth rates. Even with detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.

@@

% Could cite Blackman 1919 "The Compound Interest Law and Plant Growth" AnnBot - this approach is very old.

Earlier studies focussing on seedlings used a mathematical model to relate the trait leaf mass per area ($\phi$) to relative growth rate in mass \citep{Lambers-1992, Cornelissen-1996, Wright-2000}. In fact, that model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows. For seedlings, which are young and mostly consist of leaf biomass, we can as a first approximation ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}. Net production (from eqn \ref{eq:dbdt}) then becomes a linear function of leaf area, making relative growth rate in mass a linear function of $\phi$:
\begin{equation}\label{eq:RGR}
\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx P_{\rm net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$. Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees: namely plant height ($H$), stem-basal area ($A_{\rm st}$) or stem diameter $D$. The derivations below makes these links clear.

@@

The {\plant} adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}. We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics \citep{Huxley-1932}. Table \ref{tab:allometry} provides key equations of the model (see \citealt{Falster-2016} for full derivation). In particular, we assume that as a plant grows:
\begin{description}
\item[Assumption 1:] Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha}).
\item[Assumptions 2:] The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as}).
\item[Assumptions 3:] The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab}).
\item[Assumptions 4:] The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}).
\item[Assumption 5:] The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
\end{description}
Assumption 1 accounts for the architectural layout of the plant. Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area. To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).

@@

We now consider how trait-based trade-offs enter into the above growth model. It is essential that any trait includes both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values. For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}. In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait. The trade-offs are then implemented as follows.
\begin{description}
\item[Seed mass, $\omega$:] There is a direct energetic trade-off between a plant's fecundity and the size of its seeds. Moreover, we assume larger seeds result in larger seedlings; this seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates the inevitable energetic trade-off between growth and reproduction. We assume the fraction of energy allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ is determined by $H_{{\rm mat}}$ via a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition.
% Is this a direct biophysical consequence of varying this trait, as outlined as a necessary condition in the above?
\item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants. Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
\item[Leaf mass per unit area, $\phi$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants. Following \citet{Wright-2004}, the rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
\item[Wood density, $\rho$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants. The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly whenever $\rho$ is varied. There is very little data on rates of sapwood turnover and respiration in relation to wood density, so this latter assumption is more speculative than the equivalent assumption for leaves, which is well supported empirically.
 \end{description}

@@

\subsection{Changes in height growth rate with trait}

We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}). We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter. The first two traits considered modify behaviour primarily at the start and end of an individuals growth trajectory, and are therefore termed ``ontogenetic traits''. The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individuals growth trajectory. Across the five different traits, we observed four relatively distinct types of response. These responses are summarised in Table \ref{tab:responses} and described further below.

@@

The remaining three traits ($\nu$, $\phi$, \& $\rho$) moderate growth rates at a point along an individuals growth trajectory (Fig. \ref{fig:growth_light_height}).

@@

The results reported above and shown in Fig. \ref{fig:growth_light_height} focus on height growth rate (eqn \ref{eq:dhdt}). Corresponding results for growth rates in stem-diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs. \ref{fig:growth_light_dia}--Fig. \ref{fig:growth_light_mass} of the Supplementary Materials. In each figure, plants were grown to a suitable diameter, area, or mass. As such, changes in relative growth rates with traits show a similar patterns as absolute growth rates.

We find that for seed mass, leaf nitrogen, leaf mass per area and height at maturation, the patterns of growth rate in stem-diameter, stem-area, or above-ground mass with respect to traits mirror those observed with respect to heigh growth (Table \ref{tab:responses}). The only trait where a slightly different response was observed was for wood density. Whereas the effect of wood density on height growth tended to diminish slightly at larger sizes (Fig. \ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem-diameter, stem-area or above-ground mass. Recall that sapwood lost via turnover is turns into heartwood. Whereas the loss of sapwood diverts energy away from height growth rate, the faster accumulation of heartwood actually accelerates the growth of stem diameter and area.

@@

The model presented here extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits. Importantly, the seedling model can be derived as a special case of our more-extended approach. Unlike the original model for seedlings (eqn \ref{eq:RGR}), the new model also predicts a relationship between $\phi$ and growth rate that changes with size. In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even possibly positive in larger plants (Fig. \ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem-diameter, stem-area or total mass. This shift, which matches empirical evidence \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015, Gibert-2016}, occurs because the benefits of cheap leaf deployment diminish with plant size. As seedlings, leaves comprise a large part of the plant (Fig. \ref{fig:conceptual}a). Decreasing $\phi$ then has an overwhelmingly positive effect on height and diameter growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}. As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheaper leaf construction in eqn \ref{eq:daldmt}. Consequently, the effect of $\phi$ on leaf turnover comes to dominate at larger sizes, and as such, the effect of $\phi$ on height, diameter and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).

>>>>

% XX - The main comment I have is whether the strength of the 'correlation' between traits and growth rates can be summarized in some metric, so that it can be directly shown. To me Fig 4 is not very clear on showing how size affects the trait-growth correlations. For example, for wood density, I see a relation for all size classes. For the largest trees, it may look like a weaker correlation but isn't that because growth rates are lower? The same with Narea, to me they all look the same (except again, larger trees grow slower but it’s hard to tell there for me whether the trait-growth correlation is any different)
% XX -  The shade tolerance (WPLCP) analysis seems off-topic (it also not mentioned in the abstract). Can this be integrated more closely with the main objective, i.e. correlation between traits and growth rate and its dependence on size?
% XX - Georges suggests the graident of blues in Fig 4-5 not disntinct enough

@@

Interpreting diverse empirical results seeking to link traits to growth rate is challenging because, until the paper by \citet{Gibert-2016}, we lacked any clear expectations on why the effect of traits may be moderated by size. Generating appropriate expectations is one of the primary roles for theory \citep{Kokko-2007}. A widely-used equation for seedlings suggests that, all else being equal, a seedling's relative growth rate in mass is linearly and negatively related to LMA \citep{Lambers-1992, Cornelissen-1996, Wright-2000}. An extension of the model suggests a similar relationship should hold at larger plant sizes \citep{Enquist-2007}. But as noted above, the prediction for large plants is not supported by empirical results. Meanwhile, theoretical predictions on how other traits should influence growth are largely absent. Without further guidance, many researchers expected traits to linearly map onto growth rates, i.e. there is a ``fast'' and ``slow'' growth end for each trait \citep[e.g.][]{Grime-1977, Poorter-2008, Chave-2009, Paine-2015}.

One possible reason that theoretical predictions have either been lacking or not been supported is that the effects of traits in existing theory are realised mainly via influences on primary productivity (photosynthesis - respiration) \citep{Wright-2000, Enquist-2007}. By contrast, the physiology of traits such as LMA and wood density suggests they have influence by altering the allocation of biomass among different tissues, rather than changing mass production \citep{Falster-2011, Duursma-2016, Gibert-2016}. A second concern for theory focussed on mass production is that measuring mass production is really only practical for small plants that can be easily harvested. For larger plants, growth is more often measured as increment in either diameter or height. For example, stem diameter growth has been measured for many decades across many thousands of plots across the globe \citep{Purves-2008, Anderson-2015, Kunstler-2016}. Theory considering only primary production therefore offers limited insight.

Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence plant demography across the life-cycle. Broadly, the {\plant} model builds on several approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011}. Our primary focus in this article is to explain the pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. The paper by \citet{Gibert-2016} used the framework of the {\plant} model to make a conceptual argument why the effect of traits might change with size. Here we extend the results of \citet{Gibert-2016} to provide a full, functioning model and use this to how, from the point of view of primary production and allocation, why it makes sense that the effect of some traits changes with size. We also show how our approach can account for other phenomena, including changes in growth and shade tolerance with traits, individual size, and light environment. Our view is that trait-based approaches -- which aim to explain differences among species -- should be integrated within a general model of plant growth, and thus should also be able to capture patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{Sillett-2010, King-2011} or diameter growth \citep{Herault-2011} or biomass production \citep{Givnish-1988, Koch-2004}. In contrast, the growth rate of plant mass continues to increase with size \citep{Sillett-2010, Stephenson-2014}, and all growth measures decrease sharply with size when expressed as relative growth rates \citep{Rees-2010, Iida-2014}. Shade tolerance also varies among species, correlates with traits \citep{Messier-1999, Lusk-2008, Poorter-2006}, and tends to decrease with increasing size \citep{Givnish-1988, Kneeshaw-2006, Lusk-2008}. These diverse phenomena -- summarised in Table \ref{tab:phenomena} -- are deserving of a comprehensive and integrated explanation.

@@

The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig. \ref{fig:conceptual}a). We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive (l+b+s+r), t=total (l+b+s+h+r), and st=stem total (s+b+h). The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$. A summary of all variables, units and definitions is given in Table \ref{tab:definitions}-\ref{tab:traits}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.

We assume growth is fundamentally driven by biomass production and its subsequent distribution throughout the plant. Applying a standard approach, the amount of biomass available for growth, $\frac{{\rm d}B}{{\rm d}t}$ is given by the difference between income (photosynthesis) and losses (respiration and turnover) within the plant \citep{Makela-1997, Thornley-2000}:
\begin{equation}\label{eq:dbdt}
\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
= \underbrace{\strut \alpha_{\rm bio} }_\textrm{mass per C} \,
\underbrace{\strut \alpha_{\rm y} }_\textrm{yield}
\big( \underbrace{\strut A_{\rm l} \, \bar{p}(E)}_\textrm{photosynthesis} -
\underbrace{\strut \, \sum_{i = {\rm l}, {\rm b}, {\rm s}, {\rm r}}{M_i \, r_i}}_\textrm{respiration}\big)
- \underbrace{\strut \sum_{i = {\rm l}, {\rm b}, {\rm s},  {\rm r}}{M_i \, k_i}}_\textrm{turnover}.
\end{equation}
Photosynthesis is the product of the average photosynthetic rate per unit leaf area ($\bar{p}(E)$) and total leaf area ($A_{\rm l}$). We assume that $\bar{p}$ increases with canopy openness $E$, as per a standard light-response curve (Fig. \ref{fig:wplcp_idea}; for details, see Supplementary Materials), and respiration and turnover rates of different tissues are constants that might differ with traits (see below). The constants $\alpha_{\rm y}$ and $\alpha_{\rm bio}$ account for growth respiration and the conversion of CO$_2$ into units of biomass, respectively. While the {\plant} model can easily accommodate competitive shading via influences on $E$, in this analysis we grow individual plants under a fixed light environment, so that we can better understand the intrinsic trait- and size-related effects. We also do not yet consider any traits influencing levels of self-shading. Also, many vegetation models use a more-detailed physiological model for calculating $\bar{p}$ and $r_{\rm i}$, e.g. as functions of temperature. Such detail is not needed here because it will not change general model behaviour, even it would alter  he absolute values of predicted growth rates. Even with detailed models, $\frac{{\rm d}B}{{\rm d}t}$ always comes down to simple subtraction of income and losses.

@@

Earlier studies focussing on seedlings used a mathematical model to relate the trait leaf mass per area ($\phi$) to relative growth rate in mass \citep{Blackman-1919, Lambers-1992, Cornelissen-1996, Wright-2000}. To aid comparison with this literature, we will first show how that model is derived from eqn \ref{eq:dbdt}. In fact, the seedling model is a special case of the more-extended approach described in the following sections and can be derived from eqn \ref{eq:dbdt} as follows. For seedlings, which are young and mostly consist of leaf biomass, we can as a first approximation ignore all turnover terms as well as the respiration terms for non-leaf tissues in eqn \ref{eq:dbdt}. Net production (from eqn \ref{eq:dbdt}) then becomes a linear function of leaf area, making relative growth rate in mass a linear function of $\phi$:
\begin{equation}\label{eq:RGR}
\frac{{\rm d}B}{{\rm d}t}\frac{1}{M_{\rm a}}  \approx P_{\rm net} \times \phi^{-1} \times \frac{M_{\rm l}}{M_{\rm a}}, \end{equation}
where $P_{\rm net} = \alpha_{\rm bio} \, \alpha_{\rm y}\,(\bar{p}(E) - r_{\rm l})$. Although eqn \ref{eq:RGR} captures patterns of growth in seedlings in relation to $\phi$ \citep{Wright-2000}, this approximation does not easily extend to the variables that are routinely collected for large trees: namely plant height ($H$), stem-basal area ($A_{\rm st}$) or stem diameter $D$. The derivations below makes these links clear.

@@

The {\plant} adopts a model of plant construction and function that can be considered a first-order functional-balance or function-equilibrium model, similar to those implemented by \citet{Makela-1997} and \citet{Moorcroft-2001}. We could also call it ``isometric'', because the assumptions see area-based metrics scaling to the first-power of other area-based metrics, and to the square-power of length-based metrics, such as height \citep{Huxley-1932}. Table \ref{tab:allometry} provides key equations of the model (see \citealt{Falster-2016} for full derivation). In particular, we assume that as a plant grows:
\begin{description}
\item[Assumption 1:] Its height scales to the 0.5 power of its leaf area (eqn \ref{eq:ha}).
\item[Assumptions 2:] The cross-sectional area of sapwood in the stem is proportional to its leaf area (eqn \ref{eq:as}).
\item[Assumptions 3:] The cross-sectional area of bark and phloem in the stem  is proportional to its leaf area (eqn \ref{eq:ab}).
\item[Assumptions 4:] The cross-sectional area of root surface area and therefore mass is proportional to its leaf area (eqn \ref{eq:mr}).
\item[Assumption 5:] The vertical distribution of leaf within the plant's canopy, relative to the plant's height, remains constant.
\end{description}
Assumption 1 accounts for the architectural layout of the plant. Assumptions 2-4 are realisations of the pipe model \citep{Shinozaki-1964}, whereby the cross-sectional area of conducting tissues are proportional to leaf area. To describe the vertical distribution of leaf area within the canopy of an individual plant (assumption 5), we use the model of \citet{Yokozawa-1995}, which can account for a variety of canopy profiles through a single parameter $\eta_c$, varying from 0-1 (for details, see Supplementary Materials).

@@

We now consider how trait-based trade-offs enter into the above growth model. It is essential that any trait includes both a benefit and cost in terms of plant function and/or life history; otherwise we would expect ever-increasing trait values towards more beneficial values. For present purposes, we consider five prominent traits for which we can posit specific costs and benefits, outlined in Table \ref{tab:traits}. In postulating potential benefits and costs, we consider only those thought to arise as direct biophysical consequences of varying a trait. The trade-offs are then implemented as follows.
\begin{description}
\item[Seed mass, $\omega$:] There is a direct energetic trade-off between a plant's fecundity and the size of its seeds. Moreover, we assume larger seeds result in larger seedlings; this seed-size number tradeoff translates into a demographic tradeoff between the number of seedlings produced by a parent and their initial size.
\item[Height at maturation, $H_{{\rm mat}}$:] This trait moderates an inevitable energetic trade-off between growth and reproduction, that operates at all times through the lifestyle. Mass invested in growth cannot be invested in reproduction. To describe how the fraction of mass allocated to reproduction, $1- \frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$, changes through ontogeny, we assume a function $r(H, H_{{\rm mat}}) = \frac{r_{\rm r1}} {1.0 + \exp\left(r_{\rm r2} \left(1 - H / H_{{\rm mat}}\right)\right)}$, where $H_{{\rm mat}}$ is height at maturation, $r_{\rm r1}$ is the maximum possible allocation (0-1) and $r_{\rm r2}$ determines the sharpness of the transition. The exact shape of this function is non-critical, what is important is that plants shift from a period of investing mainly in growth to investing mainly in reproduction. The trait $H_{{\rm mat}}$ then describes the size at which this shift occurs, with direct biophysical consequences for growth.
\item[Nitrogen per unit leaf area,  $\nu$:] We allow for the maximum photosynthetic capacity of the leaf to vary with leaf nitrogen per unit area, as $A_{\rm max} = \beta_{{\rm lf1}} \, \left(\frac{\nu}{\nu_0}\right)^{\beta_{{\rm lf5}}}$, where $\beta_{{\rm lf1}}$, $\nu_0$ and $\beta_{{\rm lf5}}$ are constants. Respiration rates per unit leaf area are also assumed to vary linearly with leaf nitrogen per unit area, as $\beta_{{\rm lf4}}\, \nu$.
\item[Leaf mass per unit area, $\phi$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link $\phi$ to the rate of leaf turnover ($k_{\rm l}$), based on a widely observed scaling relationship from \citet{Wright-2004}: $k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}$ where $\beta_{{\rm kl1}}$, $\phi_0$ and $\beta_{{\rm kl2}}$ are empirical constants. Following \citet{Wright-2004}, the rate of leaf respiration per unit area is assumed independent of $\phi$, as such the mass-based rate is adjusted accordingly whenever $\phi$ is varied.
\item[Wood density, $\rho$:] This trait directly influences growth by changing $\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}$ (Table \ref{tab:allometry}). In addition, we link $\rho$ to the rate of sapwood and bark turnover, mirroring the relationship assumed for leaf turnover: $k_{\rm b} = k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^{-\beta_{{\rm ks2}}}$ where $\beta_{{\rm ks1}}$, $\rho_0$ and $\rho_{{\rm ks2}}$ are empirical constants. The rate of sapwood and bark respiration per unit stem volume is assumed to be independent of $\rho$, as such the mass-based rate is adjusted accordingly whenever $\rho$ is varied. There is very little data on rates of sapwood turnover and respiration in relation to wood density, so this latter assumption is more speculative than the equivalent assumption for leaves, which is well supported empirically.
 \end{description}

@@

\subsection{Changes in height growth rate with traits}

We analysed the response of growth rate to five different traits under the assumed trade-offs (Table \ref{tab:traits}). We considered changes in absolute and relative growth rates for mass, height, stem area and stem diameter. The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory, and are therefore termed ``ontogenetic traits''. The remaining three traits are termed ``development traits'' because they moderate the rate of movement along an individual's growth trajectory. Across the five different traits, we observed four relatively distinct types of response. These responses are summarised in Table \ref{tab:responses} and described in more detail below.

@@

The remaining three traits ($\nu$, $\phi$, \& $\rho$) moderate growth rates at a point along an individual's growth trajectory (Fig. \ref{fig:growth_light_height}).

@@

The results reported above and shown in Fig. \ref{fig:growth_light_height} focus on height growth rate (eqn \ref{eq:dhdt}). Corresponding results for growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs. \ref{fig:growth_light_dia}--Fig. \ref{fig:growth_light_mass} of the Supplementary Materials. In each figure, plants were grown to a suitable diameter, area, or mass. As such, changes in relative growth rates with traits show a similar patterns as absolute growth rates.

We find that for seed mass, leaf nitrogen, leaf mass per area and height at maturation, the patterns of growth rate in stem diameter, stem area, or above-ground mass with respect to traits mirror those observed with respect to heigh growth (Table \ref{tab:responses}). The only trait where a slightly different response was observed was for wood density. Whereas the effect of wood density on height growth tended to diminish slightly at larger sizes (Fig. \ref{fig:growth_light_height}), the effect became even stronger when measuring growth rate in stem diameter, stem area or above-ground mass. Recall that sapwood lost via turnover is turns into heartwood. Whereas the loss of sapwood diverts energy away from height growth rate, the faster accumulation of heartwood actually accelerates the growth of stem diameter and area.

@@

The model presented here extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait $\phi$ \citep{Lambers-1992, Wright-2000}, to larger plants and other traits. Importantly, the seedling model can be derived as a special case of our extended approach. Unlike the original model for seedlings (eqn \ref{eq:RGR}), the new model also predicts a relationship between $\phi$ and growth rate that changes with size. In particular, the correlation shifts from being strongly negative in seedlings to being absent, or even possibly positive in larger plants (Fig. \ref{fig:growth_light_height}), irrespective of whether growth rate is estimate via height, stem diameter, stem area or total mass. This shift, which matches empirical evidence \citep{Poorter-2008, Wright-2010, Herault-2011, Paine-2015, Gibert-2016}, occurs because the benefits of cheap leaf deployment diminish with plant size. As seedlings, leaves comprise a large part of the plant (Fig. \ref{fig:conceptual}a). Decreasing $\phi$ then has an overwhelmingly positive effect on height and diameter growth rate because the effect of increasing $\frac{\ud A_{\rm l}}{\ud M_{\rm l}}$ is large compared to the other terms in eqn \ref{eq:daldmt}. As plants increase in size, however, the amount of supporting tissue increases (Fig \ref{fig:assumptions}d), decreasing the benefit of cheaper leaf construction in eqn \ref{eq:daldmt}. Consequently, the effect of $\phi$ on leaf turnover comes to dominate at larger sizes, and as such, the effect of $\phi$ on height, diameter and, mass growth shifts from negative to either flat or mildly positive  (Table \ref{tab:responses}).
